Natural Language Interfaces to Structured Data

### 2012-08-27 trunk/tr-nlisurvey/src/nli-survey.tex
Standard commit Following sections have been updated: section 4: picture
minor

@@

\begin{figure*}
\centering
\begin{tikzpicture}[transform shape]
%\tikzset{VertexStyle/.append style = {minimum size = 3pt}}
\node at (0,0) [draw,circle,inner sep=2pt](1){1};
\node at (3,0) [draw,circle,inner sep=2pt](2){2};
\node at (7,0) [draw,circle,inner sep=2pt](3){3};
\node at (10,0) [draw,circle,inner sep=2pt](4){4};
\node at (14,0) [draw,circle,inner sep=2pt](5){5};
\node at (0,0) [draw=none,fill=none,label={[label distance=3pt]0:{NL query}}]
{}; 
\node at (3,0) [draw=none,fill=none,label={[label
distance=3pt]0:{intermediate query}}] {}; 
\node at (7,0) [draw=none,fill=none,label={[label
distance=3pt]0:{query graph}}] {};
\node at (10,0) [draw=none,fill=none,label={[label
distance=3pt]0:{logic query graph}}] {};
\node at (14,0) [draw=none,fill=none,label={[label
distance=3pt]0:{database query}}] {};
\draw[->](1) edge[bend left] (2);
\draw[->](2) edge[bend left] (3);
\draw[->](3) edge[bend left] (4);
\draw[->](4) edge[bend left] (5);
\end{tikzpicture}
\caption{Big picture of NLI systems}
\label{fig:big-picture}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ANATOMY OF SYSTEMS                  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Anatomy of NLI systems}
\label{sec:anatomy}
The main problem to solve is to map user's intent expressed in a natural way
(say unstructured way) to a database query, which is a structured expression
where there is no room for ambiguity unlike natural language. The problem that
NLI systems try to solve, is equivalent to finding a mapping $f$ between natural
language questions $q$ and families of structured queries $(q^\prime_i)_i$
(where the index corresponds to the \emph{rank} of the structured query):
\begin{equation}
f:\left\{
\begin{array}{l}
Q\rightarrow (Q^\prime)^I\\
q\mapsto(q^\prime_i)_i
\end{array}\right.
\end{equation}
where $i\in I=[0,n]$ is the index of $q^\prime_i$.

@@

\textbf{Classic translation approach}
The classic translation approach consists in going from (1) to (4).
To reach step (3), semantic rules are needed to find out what database elements
should be associated to question phrases.

\textbf{Iterative approach}
The iterative approach does not go directly from (1) to (3), but goes back to
(1) several times before reaching step (3). 
Those iterations correspond to question reformulations, that involve
user-feedback to interpret semantically the question in terms of a database
query.

>>>>

\usepackage{rotating}

@@

\begin{figure*}
\centering
\begin{tikzpicture}[transform shape]
%\tikzset{VertexStyle/.append style = {minimum size = 3pt}}
\node at (0,0) [draw,circle,inner sep=2pt](1){1};
\node at (3,0) [draw,circle,inner sep=2pt](2){2};
\node at (7,0) [draw,circle,inner sep=2pt](3){3};
\node at (10,0) [draw,circle,inner sep=2pt](4){4};
\node at (14,0) [draw,circle,inner sep=2pt](5){5};
\node at (0,0) [draw=none,fill=none,label={[label distance=3pt]0:{NL query}}]
{}; 
\node at (3,0) [draw=none,fill=none,label={[label
distance=3pt]0:{intermediate query}}] {}; 
\node at (7,0) [draw=none,fill=none,label={[label
distance=3pt]0:{query graph}}] {};
\node at (10,0) [draw=none,fill=none,label={[label
distance=3pt]0:{logic query graph}}] {};
\node at (14,0) [draw=none,fill=none,label={[label
distance=3pt]0:{database query}}] {};
\draw[->](1) edge[bend left] (2);
\draw[->](2) edge[bend left] (3);
\draw[->](3) edge[bend left] (4);
\draw[->](4) edge[bend left] (5);
\draw[<->](1) edge[bend right] (3);
\draw[->](4) edge[bend left] (3);
\end{tikzpicture}
\caption{Big picture of NLI systems}
\label{fig:big-picture}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ANATOMY OF SYSTEMS                  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Anatomy of NLI systems}
\label{sec:anatomy}
The main problem to solve is to map user's intent expressed in a natural way
(say unstructured way) to a database query, which is a structured expression
where there is no room for ambiguity unlike natural language. The problem that
NLI systems try to solve, is equivalent to finding a mapping $f$ between natural
language questions $q$ and families of structured queries $(q^\prime_i)_i$
(where the index corresponds to the \emph{rank} of the structured query):
% \begin{equation}
$$f:\left\{
\begin{array}{l}
Q\rightarrow (Q^\prime)^I\\
q\mapsto(q^\prime_i)_i
\end{array}\right.$$
% \end{equation}
where $i\in I=[0,n]$ is the index of $q^\prime_i$.

@@

Figure~\ref{fig:big-picture} depicts the big picture of translating NL queries
to database queries.

\textbf{Classic translation approach}
The classic translation approach consists in going from
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}}
to \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}} through
 \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}},
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}}
and 
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}} (step
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {2}}} being optional).

To reach step (3), semantic rules are needed to find out what database elements
should be associated to question phrases.

\textbf{Iterative approach}
The iterative approach does not go directly from
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {1}}} to
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}, but goes back to
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {3}}} (and/or to
\raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {4}}}) several times before
reaching step \raisebox{.5pt}{\textcircled{\raisebox{-.9pt} {5}}}.
Those iterations correspond to question reformulations, that involve
user-feedback to interpret semantically the question in terms of a database
query.
